Chemically Self-Consistant Modeling of the Globular Cluster NGC 2808 and its Effects on the Inferred Helium abundance of Multiple Stellar Populations.

### 2023-09-01 src/sections/fidanka.tex
Updated with actual results finally

We do not believe that any currently available software is a match for
our use case. Therefore, we elect to develop our own software suite, \fidanka.
\fidanka is a python package designed to automate much of the process of
measuring fiducial lines in CMDs, adhering to the four criteria we lay out
above. Primary features of \fidanka may be separated into three primary
categories: fiducial line measurement, stellar population synthesise, and
isochrone optimization/fitting. Additionally, there are utility functions which
are detailed in the \fidanka documentation.

@@

Maximization is preformed using the Dirichlet process, which is a
non-parametric Bayesian method of determining the number of Gaussian distributions, $K$,
which best fit the data {\color{red}[CITATION]}. Use of the Dirichlet process
allows for dynamic variation in the number of inferred populations from
magnitude bin to magnitude bin. Specifically, populations are clearly visually
separated from the lower main sequence through the turn off; however, at the
turn off and throughout much of the subgiant branch, the two visible
populations overlap due to their extremely similar ages ({\color{red}
Citation}). The Dirichlet process allows for the BGMM method to infer a single
population in these regions, while inferring two populations in regions where
they are clearly separated. More generally, the use of the Dirichlet process
removes the need for a prior on the exact number of populations to fit. Rather,
the user specifies a upper bound on the number of populations within the
cluster. An example bin (F814W = 20.6) is shown in Figure \ref{fig:BGMMDist}.

@@

\fidanka uses MIST formatted isochrones {\color{red}[CITATION]} as input along
with distance modulus, B-V color excess, binary mass fraction, and bolometric
corrections. An arbitrarily large number of isochrones may be used to define an
arbitrary number of populations. Synthetic stars are samples from each
isochrone based on a definable probability (for example it is believed that
$\sim90\%$ of stars in globular clusters are younger population
{\color{red}[CITATION]}). Based on the metallicity, $\mu$, and E(B-V) of each
isochrone, bolometric corrections are taken from bolometric correction tables.
Where bolometric correction tables do not include exact metallicities or
extinctions a linear interpolation is preformed between the two bounding
values. {\color{red}[FIGURE]} shows an example of a synthetic CMD generated
from a set of 2 NGC 2808 isochrones as well as a comparison between those
isochrones and the measured fiducial line of the synthetic population.

>>>>

We do not believe that any currently available software is a match for
our use case. Therefore, we elect to develop our own software suite, \fidanka.
\fidanka is a python package designed to automate much of the process of
measuring fiducial lines in CMDs, adhering to the four criteria we lay out
above. Primary features of \fidanka may be separated into three
categories: fiducial line measurement, stellar population synthesise, and
isochrone optimization/fitting. Additionally, there are utility functions which
are detailed in the \fidanka documentation.

@@

Maximization is preformed using the Dirichlet process, which is a
non-parametric Bayesian method of determining the number of Gaussian distributions, $K$,
which best fit the data \citep{Ferguson1973, scikit-learn}. Use of the Dirichlet process
allows for dynamic variation in the number of inferred populations from
magnitude bin to magnitude bin. Specifically, populations are clearly visually
separated from the lower main sequence through the turn off; however, at the
turn off and throughout much of the subgiant branch, the two visible
populations overlap due to their extremely similar ages \citep[i.e.][]{Jordan2002}. The Dirichlet process allows for the BGMM method to infer a single
population in these regions, while inferring two populations in regions where
they are clearly separated. More generally, the use of the Dirichlet process
removes the need for a prior on the exact number of populations to fit. Rather,
the user specifies a upper bound on the number of populations within the
cluster. An example bin (F814W = 20.6) is shown in Figure \ref{fig:BGMMDist}.

@@

\fidanka uses MIST formatted isochrones \citep{Dotter2016} as input along
with distance modulus, B-V color excess, binary mass fraction, and bolometric
corrections. An arbitrarily large number of isochrones may be used to define an
arbitrary number of populations. Synthetic stars are samples from each
isochrone based on a definable probability (for example it is believed that
$\sim90\%$ of stars in globular clusters are younger population
{\color{red}[CITATION]}). Based on the metallicity, $\mu$, and E(B-V) of each
isochrone, bolometric corrections are taken from bolometric correction tables.
Where bolometric correction tables do not include exact metallicities or
extinctions a linear interpolation is preformed between the two bounding
values. {\color{red}[FIGURE]} shows an example of a synthetic CMD generated
from a set of 2 NGC 2808 isochrones as well as a comparison between those
isochrones and the measured fiducial line of the synthetic population.

### 2023-09-01 src/sections/fitting.tex
Updated with actual results finally

\section{Isochrone Fitting}\label{sec:isoFit}
We fit pairs of isochrones to the HUGS data for NGC 2808 using \fidanka, as
descrbed in \S {\color{red}FIDANKA SECTION}. Two isochrones, one for Population
A and one for Population E are fit simultaneously. These isochrones are
constrained to have the same distance modulus, $\mu$, and color excess, E(B-V).
Moreover, we constrain the mixing length, $\alpha_{ML}$, to be constant between
any two isochrones in a set. For every isochrone in the set of combination of
which fulfilling these constraints $\mu$, $E(B-V)$, Age$_{A}$, and $Age_{B}$
are optimized to reduce the $\chi^{2}$ distance between the fiducial
line{\color{red}s} and the isochrones. Because we fit fiducial lines directly,
we do not need to consider the binary population fraction, $f_{bin}$, as a free
parameter.

{\color{red}Table \ref{tab:isoFit}} shows the results of the isochrone fitting
for NGC 2808. The best fit isochrones are shown in {\color{red}Figure
\ref{fig:isoFit}.}

{\color{red} Disscus the implications of the specific isochrones we fit.}

\subsection{ACS-HUGS Photometric Zero Point Offset}
The Hubble legacy archive photometry used in this work is calibrated to the
Vega magnitude system. However, we have found that the photometry has a
systematic offset of {\color{red}$\sim0.02$ magnitudes} in the F814W band when
compared to the same stars in the ACS survey {\color{red} [CITATION]}. The
exact cause of this offset is unknown, but it is likely due to a difference in
the photometric zero point between the two surveys. A full correction of this
offset would require a careful re-reduction of the HUGS photometry, which is
beyond the scope of this work. We instead recognize a 0.02 inherent uncertainty
in the inferred magnitude of any fit when comparing to the ACS survey. This
uncertainty is {\color{red} small when compared to the uncertainty in the
distance modulus and should not affect the conclusion of this
paper}.

{\color{red} Figure Showing the offset}

{\color{blue}The independent analysis of the Hubble Space Telescope photometry
ACS and HUGS, and an additional ground-based dataset created by Stetson et al.
(2009), of the globular cluster NGC2808 resulted in findings of a systematic
difference in the magnitude of the stars. A significant offset between the
three datasets was found. HUGS and ACS were found to have consistent data for
Vvega, however an offset of approximately 0.025 was found in the Ivega. ACS and
Stetson were found to have inconsistent differences in both Vvega and Ivega.
The three photometric studies do not match, and future studies which compare
theoretical models to the data will need to take into account these systematic
uncertainties. }

>>>>

\section{Isochrone Fitting}\label{sec:isoFit}
We fit pairs of isochrones to the HUGS data for NGC 2808 using \fidanka, as
descrbed in \S {\color{red}FIDANKA SECTION}. Two isochrones, one for Population
A and one for Population E are fit simultaneously. These isochrones are
constrained to have distance modulus, $\mu$, and color excess, E(B-V) which agree to within 1\%.
Moreover, we constrain the mixing length, $\alpha_{ML}$, to be constant between
any two isochrones in a set. For every isochrone in the set of combination of
which fulfilling these constraints $\mu$, $E(B-V)$, Age$_{A}$, and $Age_{B}$
are optimized to reduce the $\chi^{2}$ distance between the fiducial
lines and the isochrones. Because we fit fiducial lines directly,
we do not need to consider the binary population fraction, $f_{bin}$, as a free
parameter.

The best fit isochrones are shown in Figure \ref{fig:BestFitResults} and optimized
parameters for these are presented in Table \ref{tab:BestFitResults}. We find helium mas fractions which are consistent with those identified in past literature \citep[e.g.][]{Milone2015}. Note that our helium mass fraction gird has a spacing of 0.03 between grid points and we are therefore unable to resolve between certain proposed helium mass fractions for the younger sequence (for example between 0.37 and 0.39).

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{src/figures/BestFitResults.pdf}
  \label{fig:BestFitResults}
  \caption{Best fit isochrone results for NGC 2808.}
\end{figure}

\begin{table*}
  \centering
  \begin{tabular}{c | c c c c c c}
    \hline
    population & age & distance modulus & extinction & Y & $\alpha_{ML}$ & $\chi^{2}_{\nu}$\\
    & [Gyr] & & [mag] & & &\\
    \hline
    \hline
    A & 12.3 & 14.91 & 0.54 & 0.24 & 1.901 & 0.014\\
    E & 14.3 & 14.96 & 0.54 & 0.39 & 1.750 & 0.017 \\
    \hline
  \end{tabular}
  \label{tab:BestFitResults}
  \caption{Best fit parameters derived from fitting isochrones to the fiducual lines derived from the NCG 2808 photometry.}
\end{table*}

\subsection{The Number of Populartions in NGC 2808}
\fidanka provides a somewhat straigtforward way to estimate the number of populations expected in a given magnitude bin given the observations. See Section \ref{sec:fidanka} for specific implimentaiton details. Here we preform an analysis of the number of populations seen in the NGC 2808 F814W-F274W vs F814W color-magnitude diagram. We find that for the majority of the main sequence and red giant branches BGMM prefers two populations; wherease, near the main sequnce turn off and on the majority of the subgiant branches BGMM prefers a single population model.

{\color{red} Make Figure showing the BGMM fit predictions for each star vs. distance to pop mean line, does a good job of visualizing the predicted number of populations}

\subsection{ACS-HUGS Photometric Zero Point Offset}
The Hubble legacy archive photometry used in this work is calibrated to the
Vega magnitude system. However, we have found that the photometry has a
systematic offset of $\sim0.026$ magnitudes in the F814W band when
compared to the same stars in the ACS survey (Figure \ref{fig:offset}). The
exact cause of this offset is unknown, but it is likely due to a difference in
the photometric zero point between the two surveys. A full correction of this
offset would require a careful re-reduction of the HUGS photometry, which is
beyond the scope of this work. We instead recognize a 0.02 inherent uncertainty
in the inferred magnitude of any fit when comparing to the ACS survey. This
uncertainty is small when compared to the uncertainty in the
distance modulus and should not affect the conclusion of this
paper.

\begin{figure*}
  \centering
  \includegraphics[width=0.90\textwidth]{src/figures/photometricOffset.pdf}
  \label{fig:offset}
  \caption{(left) CMD showing the photometric offset between the ACS and HUGS data for NGC 2808. CMDs have been randomly subsampled and colored by point density for clarity. (right) Mean difference between the color of the HUGS and ACS fiducual lines at the same magnitude. Note that the ACS data is systematically bluer than the HUGS data.}
\end{figure*}

The oberved photometric offset between ACS and HUGS reductions introduces a systematic uncertainity when comparing parameters derived from isochrone fits to ACS data vs those fit to HUGS data. Specifically, this offset introduces a {\color{red}$\sim$AGE Gyr} uncertainity. Moreover, for two isochrone of the same age, only seperated by helium mass fraction, a shift of the main sequence turn off of is also expected (Figure {\color{red} Figure showing MSTO location for all Ys}). {\color{red}Compare this to size of the MSTO shift seen here and how that informs the Y uncertainity we work with when comparing to ACS data.}

### 2023-09-01 src/sections/intro.tex
Updated with actual results finally

The traditional view of Globular Clusters was, for a long time, that they
consisted of a single stellar population (SSP, in some publications this is
referred to as a Simple Stellar Population). This view was supported by
spectroscopically uniform heavy element abundances \citep{Carretta2009}
{\color{red}[CHECK]}, and the lack of evidence for multiple stellar populations
(MPs) in the color-magnitude diagrams of GCs \citep{Milone2017}. However, in
the last two decades, it has been definitively shown that most if not all Milky
Way GCs have MPs \citep{Piotto2015}. The lack of photometric evidence for MPs
can be attributed to the short color throw available to ground based
photometric surveys \citep{Milone20Here17}. While MPs are chemically distinct
from one another, that distinction is most prominent when observing with $U$
and $B$ filters \citep{Sbordone2011}.

The prevalence of multiple populations in GCs is so distinct that the proposed
definitions for what constitutes a globular cluster now often center the
existence of MPs. Whereas, people have have often tried to categorized objects
as GCs through relations between half-light radius, density, and surface
brightness profile, in fact many objects which are generally thought of as GCs
don't cleanly fit into these cuts {\color{red} EXAMPLE + CITATION}.
Consequently, \citet{Carretta2010} proposed a definition of GC based on
observed chemical inhomogeneities in their stellar populations. The modern
understanding of GCs then is not simply one of a dense cluster of stars which
may have chemical inhomogeneities and multiple populations; rather, it is one
where those chemical inhomogeneities and multiple populations themselves are
the defining element of a GC.

>>>>

The traditional view of Globular Clusters was, for a long time, that they
consisted of a single stellar population (SSP, in some publications this is
referred to as a Simple Stellar Population). This view was supported by
spectroscopically uniform heavy element abundances \citep{Carretta2010, Bastian2018} accross most clusters (M54 and $\omega$Cen are notable exceptions, see \citet{Marino2015} for further details), and the lack of evidence for multiple stellar populations
(MPs) in past color-magnitude diagrams of GCs \citep[i.e.][]{Sandage1953, Alcaino1975}. However, over the last 40 years non-trivial star-to-star light-element abundance variations have been observed \citep[i.e.][]{Smith1987} and, in
the last two decades, it has been definitively shown that most if not all Milky
Way GCs have MPs \citep{Gratton2004, Gratton2012, Piotto2015}. The lack of photometric evidence for MPs
can be attributed to the short color throw available to ground based
photometric surveys \citep{Milone2017}; specifically, lacking UV filters. While MPs are chemically distinct
from one another, that distinction is most prominent when observing with $U$
and $B$ filters \citep{Sbordone2011}.

The prevalence of multiple populations in GCs is so distinct that the proposed
definitions for what constitutes a globular cluster now often center the
existence of MPs. Whereas, people have have often tried to categorized objects
as GCs through relations between half-light radius, density, and surface
brightness profile, in fact many objects which are generally thought of as GCs
don't cleanly fit into these cuts \citep{Peebles1968, Brown1991, Brown1995, Bekki2002}.
Consequently, \citet{Carretta2010} proposed a definition of GC based on
observed chemical inhomogeneities in their stellar populations. The modern
understanding of GCs then is not simply one of a dense cluster of stars which
may have chemical inhomogeneities and multiple populations; rather, it is one
where those chemical inhomogeneities and multiple populations themselves are
the defining element of a GC.

### 2023-09-01 src/sections/modeling.tex
Added some more placeholder sections
Updated with actual results finally

We use DSEP to evolve stellar models ranging in mass from 0.3 to 2.0 solar
masses from the zero-age main sequence (ZAMS) to the tip of the red giant
branch. Below 0.7 $M_{\odot}$ we evolve a model every 0.03 $M_{\odot}$ and
above 0.7 $M_{\odot}$ we evolve a model every 0.5 $M_{\odot}$. Additionally, we
evolve models over a grid of mixing length parameters, $\alpha_{MLT}$, from
$\alpha_{MLT} = 1.0$ to $\alpha_{MLT} = 2.0$ in steps of 0.1. In addition to the
mixing length grid the evolved grid of models also has dimensions population (A
or E) and helium abundance ($Y$). Each model is evolved in DSEP's ``high
resolution'' mode and had a maximum allowed time step of 50 Myr.

>>>>

We use DSEP to evolve stellar models ranging in mass from 0.3 to 2.0 solar
masses from the zero-age main sequence (ZAMS) to the tip of the red giant
branch. Below 0.7 $M_{\odot}$ we evolve a model every 0.03 $M_{\odot}$ and
above 0.7 $M_{\odot}$ we evolve a model every 0.5 $M_{\odot}$. Additionally, we
evolve models over a grid of mixing length parameters, $\alpha_{MLT}$, from
$\alpha_{MLT} = 1.0$ to $\alpha_{MLT} = 2.0$ in steps of 0.1. In addition to the
mixing length grid the evolved grid of models also has dimensions population (A
or E) (Table \ref{tab:comp}) and helium abundance ($Y=0.24, 0.27, 0.3, 0.33, 0.36, 0.39$). Each model is evolved in DSEP's ``high
resolution'' mode and had a maximum allowed time step of 50 Myr.

\begin{deluxetable}{c|cc||c|cc}\label{tab:comp}

%% Keep a portrait orientation

%% Over-ride the default font size
%% Use Default (12pt)

%% Use \tablewidth{?pt} to over-ride the default table width.
%% If you are unhappy with the default look at the end of the
%% *.log file to see what the default was set at before adjusting
%% this value.

%% This is the title of the table.
\tablecaption{Population Composition}

%% This command over-rides LaTeX's natural table count
%% and replaces it with this number.  LaTeX will increment 
%% all other tables after this table based on this number
\tablenum{1}

%% The \tablehead gives provides the column headers.  It
%% is currently set up so that the column labels are on the
%% top line and the units surrounded by ()s are in the 
%% bottom line.  You may add more header information by writing
%% another line between these lines. For each column that requries
%% extra information be sure to include a \colhead{text} command
%% and remember to end any extra lines with \\ and include the 
%% correct number of &s.
  \tablehead{\colhead{Element} & \colhead{Pop A} & \colhead{Pop E} & \colhead{Element} & \colhead{Pop A} & \colhead{Pop E} 
}

%% All data must appear between the \startdata and \enddata commands
\startdata
Li & -0.08 & --- & In & -1.46 & --- \\
Be & 0.25 & --- & Sn & -0.22 & --- \\
B & 1.57 & --- & Sb & -1.25 & --- \\
C & 6.87 & 5.91 & Te & -0.08 & --- \\
N & 6.42 & 6.69 & I & -0.71 & --- \\
O & 7.87 & 6.91 & Xe & -0.02 & --- \\
F & 3.43 & --- & Cs & -1.18 & --- \\
Ne & 7.12 & 6.7 & Ba & 1.05 & --- \\
Na & 5.11 & 5.7 & La & -0.03 & --- \\
Mg & 6.86 & 6.42 & Ce & 0.45 & --- \\
Al & 5.21 & 6.61 & Pr & -1.54 & --- \\
Si & 6.65 & 6.77 & Nd & 0.29 & --- \\
P & 4.28 & --- & Pm & -99.0 & --- \\
S & 6.31 & 5.89 & Sm & -1.3 & --- \\
Cl & -1.13 & 4.37 & Eu & -0.61 & --- \\
Ar & 5.59 & 5.17 & Gd & -1.19 & --- \\
K & 3.9 & --- & Tb & -1.96 & --- \\
Ca & 5.21 & --- & Dy & -1.16 & --- \\
Sc & 2.02 & --- & Ho & -1.78 & --- \\
Ti & 3.82 & --- & Er & -1.34 & --- \\
V & 2.8 & --- & Tm & -2.16 & --- \\
Cr & 4.51 & --- & Yb & -1.42 & --- \\
Mn & 4.3 & --- & Lu & -2.16 & --- \\
Fe & 6.37 & --- & Hf & -1.41 & --- \\
Co & 3.86 & --- & Ta & -2.38 & --- \\
Ni & 5.09 & --- & W & -1.41 & --- \\
Cu & 3.06 & --- & Re & -2.0 & --- \\
Zn & 2.3 & --- & Os & -0.86 & --- \\
Ga & 0.78 & --- & Ir & -0.88 & --- \\
Ge & 1.39 & --- & Pt & -0.64 & --- \\
As & 0.04 & --- & Au & -1.34 & --- \\
Se & 1.08 & --- & Hg & -1.09 & --- \\
Br & 0.28 & --- & Tl & -1.36 & --- \\
Kr & 0.99 & --- & Pb & -0.51 & --- \\
Rb & 0.26 & --- & Bi & -1.61 & --- \\
Sr & 0.61 & --- & Po & -99.0 & --- \\
Y & 1.08 & --- & At & -99.0 & --- \\
Zr & 1.45 & --- & Rn & -99.0 & --- \\
Nb & -0.8 & --- & Fr & -99.0 & --- \\
Mo & -0.38 & --- & Ra & -99.0 & --- \\
Tc & -99.0 & --- & Ac & -99.0 & --- \\
Ru & -0.51 & --- & Th & -2.2 & --- \\
Rh & -1.35 & --- & Pa & -99.0 & --- \\
Pd & -0.69 & --- & U & -2.8 & --- \\
\enddata

%% Include any \tablenotetext{key}{text}, \tablerefs{ref list},
%% or \tablecomments{text} between the \enddata and 
%% \end{deluxetable} commands

%% General table comment marker
\tablecomments{Relative Metal composition used where a(H) = 12. Where the relative composition is the the same for both populations A and E it is only listed in the population A colum for the sake of visual clarity.}

%% General table references marker
\tablerefs{\citet{Milone2015}}

\end{deluxetable}

### 2023-09-01 src/sections/results.tex
Updated with actual results finally

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{src/figures/BestFit.png}
  \label{fig:BestFitResults}
  \caption{Best fit isochrone results for NGC 2808.}
\end{figure}

@@

>>>>

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{src/figures/BestFitResults.pdf}
  \label{fig:BestFitResults}
  \caption{Best fit isochrone results for NGC 2808.}
\end{figure}

\begin{table*}
  \centering
  \begin{tabular}{c | c c c c c c}
    \hline
    population & age & distance modulus & extinction & Y & $\alpha_{ML}$ & $\chi^{2}_{\nu}$\\
    & [Gyr] & & mag & & &\\
    \hline
    \hline
    A & 12.3 & 14.91 & 0.54 & 0.24 & 1.901 & 0.014\\
    E & 14.3 & 14.96 & 0.54 & 0.39 & 1.750 & 0.017 \\
    \hline
  \end{tabular}
  \label{tab:BestFitResults}
  \caption{Best fit parameters derived from fitting isochrones to the fiducual lines derived from the NCG 2808 photometry.}
\end{table*}

@@

\subsection{The Number of Populartions in NGC 2808}
\fidanka provides a somewhat straigtforward way to estimate the number of populations expected in a given magnitude bin given the observations. See Section \ref{sec:fidanka} for specific implimentaiton details. Here we preform an analysis of the number of populations seen in the NGC 2808 F814W-F274W vs F814W color-magnitude diagram. We find that for the majority of the main sequence and red giant branches BGMM prefers two populations; wherease, near the main sequnce turn off and on the majority of the subgiant branches BGMM prefers a single population model.

{\color{red}[FIGURE SHOWING BGMM population probability]}
